Characterizing Signal Loss, Error, and Bias in the 21cm Reionization \\Power Spectrum: A Revised Study of PAPER-64

### 2018-04-11 PAPER.tex
Tue Apr 10 22:22:23 EDT 2018

@@

\subsubsection{Toy Model: Inverse Covariance Weighting}
\label{sec:toymodel}

@@

In practice, we do not have perfect models for $\textbf{C}$, and it is the use of an estimated covariance $\widehat{\textbf{C}}$ instead of the true covariance $\textbf{C}$ that can lead to loss. We will now build intuition into how empirically estimated inverse covariance weighting can give rise to signal loss through the use of a toy model.

We construct a simple dataset that contains visibility data with $100$ time 
integrations and $20$ frequency channels. This model represents realistic dimensions of about an hour of PAPER data which 
might be used for a power spectrum analysis. For PAPER-64 (both the \citetalias{ali_et_al2015} analysis and our new analysis) we use $\sim8$ hours of data (with channel 
widths of $0.5$ MHz and integration times of $43$ seconds), but here we scale it down with no loss of 
generality.

@@

@@

To begin to understand the lossy behavior of this result, we can closely study our estimated covariance eigenspectrum, shown in Figure \ref{fig:toy_sigloss2}. Since it is estimated from our data, its eigenspectrum differs from the eigenspectrum of the true covariance $\textbf{C}$, and this difference has important consequences on our result.

An eigenspectrum ranks the eigenvalues of a matrix from 
highest to lowest and can be thought of as a spectrum of weights that are given to each spectral mode in the data. In other 
words, the eigenvalues encode the strength of different shapes in the dataset. The eigenspectrum of the identity matrix $
\textbf{I}$ is flat (all $1$'s) because it gives equal weighting to all modes. In our application, covariance matrices tend to have sloped eigenspectra, meaning that modes are given different weights in QE power spectrum estimation. The modes with the highest eigenvalues are 
down-weighted the most.

@@

Because $\widehat{\textbf{C}}$ is estimated from the data, its eigenvectors and eigenvalues are strongly coupled to the particular data realization that was used to compute it. For example, the highest-valued eigenmode of $\widehat{\textbf{C}}$ (highest eigenvalue) describes the sinusoid foreground mode in the toy model (the peak in Figure \ref{fig:toy_sigloss3}). In Figure \ref{fig:toy_sigloss12} we show the estimated covariances of our toy model datasets along with $\widehat{\textbf{C}}^{-1}$ weighted data. The foreground sinusoid is clearly visible in $\widehat{\textbf{C}}_{\rm FG}$.

In general, the highest-valued eigenmodes of $\widehat{\textbf{C}}$ typically (but not always) describe bright foregrounds --- the most prominent shapes in a dataset. For these FG-dominating modes, where foregrounds outshine the EoR signal, down-weighting is beneficial. In some sense, we desire signal loss in this regime, if by `signal' we mean `foregrounds.' In this case it is beneficial for these eigenmodes to be coupled to the data in a way that produces loss. For our toy model, the successful suppression of the foreground mode is demonstrated in Figure \ref{fig:toy_sigloss3} by the missing foreground peak in the weighted power spectrum estimate (green).

>>>>

% Put all of the (O)QE / ICW theory here, rather than sprinkled throughout this section.

% Some things we never quite say as explicitly as I'd like:
% - "foreground suppression" happens because the foregrounds have a different covariance structure (Figure 4), and we 
% hope we pick this out by picking big eigenvalues
% - additive bias is a thing

\subsubsection{The Quadratic Estimator Method}

@@

@@

%In practice, we do not have perfect models for $\textbf{C}$, and it is the use of an estimated covariance $\widehat{\textbf{C}}$ instead of the true covariance $\textbf{C}$ that can lead to loss.

@@

% ---

\subsubsection{Toy Model: Empirical Inverse Covariance Weighting}
\label{sec:toymodel}

We will now build intuition into how empirically estimated inverse covariance weighting can give rise to signal loss through the use of a toy model.  We construct a simple dataset that contains visibility data with $100$ time 
integrations and $20$ frequency channels. This model represents realistic dimensions of about an hour of PAPER data which 
might be used for a power spectrum analysis. For PAPER-64 (both the \citetalias{ali_et_al2015} analysis and our new analysis) we use $\sim8$ hours of data (with channel 
widths of $0.5$ MHz and integration times of $43$ seconds), but here we scale it down with no loss of 
generality.

@@

% While a later section does show that the large eigenvalue modes converge to the eigenvectors quicker, it is _not_ true that modes with big eigenvalues are downweighted more.  It is the modes which have a structure different from dC/dp_alpha.  Our regularization actually forces both eigenvalues and eigenmodes to be like the "signal" part of the covariance.  Thus, I don't think the next two paragraphs are helping the explanation of what is going on.

%To begin to understand the lossy behavior of this result, we can closely study our estimated covariance eigenspectrum, shown in Figure \ref{fig:toy_sigloss2}. Since it is estimated from our data, its eigenspectrum differs from the eigenspectrum of the true covariance $\textbf{C}$, and this difference has important consequences on our result.

%An eigenspectrum ranks the eigenvalues of a matrix from highest to lowest and can be thought of as a spectrum of weights that are given to each spectral mode in the data. In other words, the eigenvalues encode the strength of different shapes in the dataset. The eigenspectrum of the identity matrix $ \textbf{I}$ is flat (all $1$'s) because it gives equal weighting to all modes. In our application, covariance matrices tend to have sloped eigenspectra, meaning that modes are given different weights in QE power spectrum estimation. The modes with the highest eigenvalues are down-weighted the most.

@@

Because $\widehat{\textbf{C}}$ is estimated from the data, its eigenvectors and eigenvalues are strongly coupled to the particular data realization that was used to compute it.  In this case, the eigenmode of $\widehat{\textbf{C}}$ with the largest eigenvalue describes the sinusoid foreground mode in the toy model (the peak in Figure \ref{fig:toy_sigloss3}).  In Figure \ref{fig:toy_sigloss12} we show the estimated covariances of our toy model datasets along with $\widehat{\textbf{C}}^{-1}$ weighted data. The foreground sinusoid is clearly visible in $\widehat{\textbf{C}}_{\rm FG}$.

%In general, 
We have made the assumption that the highest-valued eigenmodes of $\widehat{\textbf{C}}$ will typically (but not always) describe bright foregrounds, though we have not checked that indeed the eigenvectors do describe correlations between frequencies.
% --- the most prominent shapes in a dataset. 
%For these FG-dominating modes, where foregrounds outshine the EoR signal, 
If such modes do describe foregrounds, then down-weighting is beneficial. In some sense, we desire signal loss in this regime, if by `signal' we mean `foregrounds.'  In this case it is beneficial for these eigenmodes to be coupled to the data in a way that produces loss. For our toy model, the successful suppression of the foreground mode is demonstrated in Figure \ref{fig:toy_sigloss3} by the missing foreground peak in the weighted power spectrum estimate (green).
